---
title: "Chapter 4 - R Notation"
author: "Aodhagan"
date: "1/6/2020"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# load the data deck from website
```{r test}
rm(list = ls())

# partition the URL to make it more readable
url_remote <- "https://gist.githubusercontent.com/"
path_github <- "garrettgman/9629323/raw/ee5dfc039fd581cb467cc69c226ea2524913c3d8/"
filename <- "deck.csv"

# assign full URL to variable file_path
file_path <- paste0(url_remote, path_github, filename)

# read in the file from website
library(data.table)
deck <- fread(file_path, header = TRUE)
```

# selecting values
you can extract the value or set of values from a data frame using indexing with [ , ]
```{r}
deck[1, 1] # first row, first column
```

## six methods of indexing {.tabset}
### 1. Positive integers
e.g.
```{r}
deck[1, 2]
```
use vectors to select more than one item
```{r}
deck[1, c(1, 2, 3)]
```
**note** that the methods used for indexing data frames can also be used for other R objects such as Vectors, matrices or arrays

**note** indexing in R begins at 1

**note** if two or more columns are selected in the index a data frame is returned. 
If only one column is selected a vector is returned.
```{r}
deck[1:3, 2]
```
if you want to select one column and to have a data frame returned use the argument **drop=FALSE**
```{r}
df <- deck[1:2, 1, drop=FALSE]
is.data.frame(df)
```


### 2. negative integers
negative integers to the exact opposite of positive integers when indexing i.e. every element except the elements in the negative index will be returned
```{r}
# only the first and second rows will be indexed
deck[-(3:52), 1:3]
```
negative integers are more efficient when you want a subset the majority of data frames rows or columns

### 3. zeroes
R will return nothing from a dimension when you use zero as an index
```{r}
deck[0, 0] # data frame with zero columns and zero rows
```

### 4. Blank spaces
blank space extracts every value in a dimension
```{r}
deck[1, ] # extracts the entire first row of data frame
```

### 5. Logical values
you can index a R object using vectors of logical values
```{r}
deck[1, c(TRUE, TRUE, FALSE)]
```
every index where the vector states "TRUE" will return the corresponding index in the data frame

### 6. names
you can specify elements in a data frame by name as long as the name attributes exist 
```{r}
deck[1, c("face", "suit", "value")]
```

## exercise 1
```{r}
deal <- function(cards){
  cards[1, ]
}
deal(deck)
```

# shuffle the deck
Begin by extracting every row in the data frame
```{r}
deck2 <- deck[1:52, ]
head(deck2)
```
note that above the order hasn't changed at all. We can get R to extract the rows in a different order
```{r}
deck3 <- deck[c(2, 1, 3:52), ]
head(deck3)
```
how do we apply the same principle to get a random order?use the randomising function sample()
```{r}
random <- sample(1:52, size = 52)
deck4 <- deck[random, ]
head(deck4)
```
## exercise - create a shuffle function
```{r}
shuffle <- function(cards){
  random <- sample(1:52, size = 52)
  cards[random, ]
}
```
now we can shuffle cards between each deal
```{r}
deal(deck)

deck2 <- shuffle(deck)
deal(deck2)
```


# Dollar signs and double brackets
what about the two types of R objects that obey optional second system of notation for indexing?

```{r}
# data frames and lists can use the $ syntax
deck$value
```
what does the dollar notation of indexing return?
```{r}
#it returns a vector
mean(deck$value)
median(deck$value)
```
when subseting a list using the square bracket notation what R object is returned?
```{r}
# another list
lst <- list(numbers = c(1, 2), logical = TRUE, strings = c("a", "b", "c"))
lst
lst[1]
```
why is this disadvantageous?
```{r}
# because many functions do not work with lists
```
what R object is returned when you use the dollar notation?
```{r}
 #the selected values are returned as they are with no list structure around them
lst$numbers
```
what is the advantage of the dollar notation of indexing lists?
```{r}
# you can feed the output of an index directly into another function
sum(lst$numbers)
```

when indexing lists what notation can be used if list elements do not have names?
```{r}
# use double bracket notation
lst[[1]]
```
so what is the difference between single bracket and double bracket notation when indexing lists?
```{r}
# single brackets return a list while double brackets return just the values that were inside an element of the list
lst["numbers"]
lst[["numbers"]]
```
what does the attach() function do?
```{r}
# it allows you to refer to each variable in a data frame by name e.g.face rather than deck$face
```
why should this function not be used?
```{r}
# it recreates an environment similar to those used in other statistics applications e.g. SPSS.
# however R is optimised to use the R computing environment
```









