---
title: "Chapter 4 - R Notation"
author: "Aodhagan"
date: "1/6/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# load the data deck from website
```{r}
rm(list = ls())

# partition the URL to make it more readable
url_remote <- "https://gist.githubusercontent.com/"
path_github <- "garrettgman/9629323/raw/ee5dfc039fd581cb467cc69c226ea2524913c3d8/"
filename <- "deck.csv"

# assign full URL to variable file_path
file_path <- paste0(url_remote, path_github, filename)

# read in the file from website
library(data.table)
deck <- fread(file_path, header = TRUE)
```

# selecting values
you can extract the value or set of values from a data frame using indexing with [ , ]
```{r}
deck[1, 1] # first row, first column
```

## six methods of indexing
### 1. Positive integers
e.g.
```{r}
deck[1, 2]
```
use vectors to select more than one item
```{r}
deck[1, c(1, 2, 3)]
```
**note** that the methods used for indexing data frames can also be used for other R objects such as Vectors, matrices or arrays

**note** indexing in R begins at 1

**note** if two or more columns are selected in the index a data frame is returned. 
If only one column is selected a vector is returned.
```{r}
deck[1:3, 2]
```
if you want to select one column and to have a data frame returned use the argument **drop=FALSE**
```{r}
df <- deck[1:2, 1, drop=FALSE]
is.data.frame(df)
```


### 2. negative integers
negative integers to the exact opposite of positive integers when indexing i.e. every element except the elements in the negative index will be returned
```{r}
# only the first and second rows will be indexed
deck[-(3:52), 1:3]
```
negative integers are more efficient when you want a subset the majority of data frames rows or columns

### 3. zeroes
R will return nothing from a dimension when you use zero as an index
```{r}
deck[0, 0] # data frame with zero columns and zero rows
```

### 4. Blank spaces
blank space extracts every value in a dimension
```{r}
deck[1, ] # extracts the entire first row of data frame
```

### 5. Logical values
you can index a R object using vectors of logical values
```{r}
deck[1, c(TRUE, TRUE, FALSE)]
```
every index where the vector states "TRUE" will return the corresponding index in the data frame

### 6. names
you can specify elements in a data frame by name as long as the name attributes exist 
```{r}
deck[1, c("face", "suit", "value")]
```

## exercise 1
```{r}
deal <- function(cards){
  cards[1, ]
}
deal(deck)
```





